---
title: "Visualize and compare sgRNA abundance distributions"
subtitle: "Mouse libray (n = 2215)"
author: "Francesc Castro-Giner"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params:
  data_dir: ./data/crispr/mm_2215_sgRNA
  #date: '`r format(Sys.Date(), "%B %d, %Y")`'
  #rawdata_dir: ./data/crispr
  #data_dir: ./data/crispr/mm_2215_sgRNA
  #output_dir: ./output/crispr/mm_2215_sgRNA/dist_test
  # ncpu: 4#
---

## Load libraries, additional functions and data

Setup environment
```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(results='asis', echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, fig.align = 'center', fig.width = 3.5, fig.asp = 0.618, dpi = 600, dev = c("png", "pdf"), engine.opts = list(bash = "-l"))

options(stringsAsFactors = FALSE)

use_seed <- 1100101

set.seed(use_seed)

# dir.create(params$output_dir, recursive = TRUE, showWarnings = FALSE)
```

Load packages
```{r load-libraries}
library(tidyverse)
library(knitr)
library(foreach)
library(magrittr)
library(DT)
library(kableExtra)
library(diptest)
library(SummarizedExperiment)

library(ggridges)
library(ggh4x)
library(patchwork)
library(colorblindr)
# library(showtext)

# library(DT)
# library(colorblindr)
# library(ggdendro)
# library(RColorBrewer)
# library(circlize)
# library(Hmisc)
# library(ComplexHeatmap)
# library(ineq)

# library(kableExtra)
# library(magrittr)
# library(ggrepel)
# library(ggpubr)
# library(ggbeeswarm)
# library(ggridges)
```

Load ggplot theme
```{r ggplot-theme, include=TRUE}
source("./configuration/rmarkdown/ggplot_theme.R")
```

Load Summarized Experiment object
```{r load-se}
se <- readRDS(file.path(params$data_dir, 'se.rds'))
```

## Configure analyses

Define comparisons for modality and Kolmogorov-Smirnov test
```{r define-comparisons-modality-ks}
x <- colData(se) %>% data.frame

group_list_modality <- list(
  `Syngeneic Thy1` = x %>% 
    filter(vector_mmodel == '4T1-noKO-Thy1-BALB' & sample_type == 'whole_tumor') %>% 
    pull(sample_alias),
  
  `Syngeneic Puro` = x %>% 
    filter(vector_mmodel %in% c('4T1-noKO-Puro-BALB', 'MVT1-noKO-Puro-FVB', 'Py2T-noKO-Puro-FVB') & sample_type == 'whole_tumor') %>%
    pull(sample_alias),
  
  `Syngeneic dCas9 + Puro` = x %>% 
    filter(vector_mmodel == '4T1-noKO-Puro_EF1alpha_dCas9-BALB' & sample_type == 'whole_tumor') %>% 
    pull(sample_alias),
  
  `NSG Thy1` =  x %>% 
    filter(vector_mmodel %in% c('4T1-noKO-Thy1-NSG', 'MVT1-noKO-Thy1-NSG', 'Py2T-noKO-Thy1-NSG') & sample_type == 'whole_tumor') %>%
    pull(sample_alias),
  
  `NSG Puro` = x %>% 
    filter(vector_mmodel %in% c('4T1-noKO-Puro-NSG', 'MVT1-noKO-Puro-NSG', 'Py2T-noKO-Puro-NSG') & sample_type == 'whole_tumor') %>%
    pull(sample_alias),
  
  `NSG dCas9 + Puro` = x %>% 
    filter(vector_mmodel == '4T1-noKO-Puro_EF1alpha_dCas9-NSG' & sample_type == 'whole_tumor') %>% 
    pull(sample_alias)
)

comp_list_modality <- list(
  `Syngeneic Thy1 vs. dCas9 + Puro` = list(
    `dCas9 + Puro` = group_list_modality$`Syngeneic dCas9 + Puro`,
    `Thy1` = group_list_modality$`Syngeneic Thy1`
  ),
  `Syngeneic Thy1 vs. Puro` = list(
    `Puro` = group_list_modality$`Syngeneic Puro`,
    `Thy1` = group_list_modality$`Syngeneic Thy1`
  )
)

group_df_modality <- foreach(i = names(group_list_modality), .combine = rbind) %do% {
  data.frame(
    sample_alias = group_list_modality[[i]],
    modality_group = i
  )
}
```


## Primary tumor
```{r test-dist-tumor-conf}
use_sample_type <- 'whole_tumor'
# comp_list <- comp_list_by_sample_type
# comp_list_unlisted <- foreach(x = comp_list, .combine = c) %do% x
# comp_list_df_double <- foreach(x = comp_list_unlisted, .combine = rbind) %do% {rbind(x, rev(x))} %>% 
#   data.frame %>% 
#   set_names(c('group1', 'group2'))
use_colData <- colData(se) %>% data.frame %>%
  filter(sample_type == use_sample_type) %>% 
  mutate(
    vector_short = ifelse(vector_short == 'Puro_EF1alpha_dCas9', 
                          'dCas9+Puro', vector_short),
    vector_short = factor(vector_short, levels = c('Thy1', 'Puro', 'dCas9+Puro')),
    mouse_model_short = case_match(mouse_model_global,
                                   'immunodeficient' ~ 'NSG',
                                   'immunocompetent' ~ 'Syngeneic'
                                   ),
    mouse_model_short = factor(mouse_model_short, levels = c('NSG', 'Syngeneic'))
  )
```

### Figure 2B: sgRNA distribution in mice transplanted with 4T1 cells

Density plots showing the distribution of unique sgRNAs retrieved in orthotopic primary tumors from the indicated mice transplanted with 4T1 cells, previously transduced as indicated; frequency is shown as read counts per million (cpm) in log10 scale adding a pseudocunt of 1; Each plot is representative of a group of n=5 mice. 

```{r test-dist-tumor-distribution-ridges_plot-4t1, fig.width=3, fig.asp=1}
selected_panels <- list(
  `Thy1, NSG` = c('4T_TumoT_N4', '4T_TumoT_N5', '4T_TumoT_N2'),
  `Thy1, Syngeneic` = c('4T_TumoT_B1', '4T_TumoT_B3', '4T_TumoT_B4'),
 
  `Puro, NSG` = c('4T_TumoP_N5', '4T_TumoP_N4', '4T_TumoP_N2'),
  `Puro, Syngeneic` = c('4T_TumoP_B3', '4T_TumoP_B2', '4T_TumoP_B1'),
  
  `dCas9+Puro, NSG` = c('PlatePositionE1', 'PlatePositionC1', 'PlatePositionA2'),
  `dCas9+Puro, Syngeneic` = c('PlatePositionE2', 'PlatePositionG2', 'PlatePositionC2')
  
)

selected_panels_df <- foreach(i = names(selected_panels), .combine = rbind) %do% {
  data.frame(group = i, sample_alias = selected_panels[[i]])
}

use_samples <- unlist(selected_panels)
use_assay <- assay(se[,use_samples], 'cpm') %>% 
  as.data.frame(check.names = F) %>% 
  rownames_to_column('guide') %>% 
  pivot_longer(-guide, names_to = 'sample_alias', values_to = 'cpm') %>% 
  left_join(use_colData) %>% 
  left_join(selected_panels_df) %>% 
  mutate(
    sample_alias = factor(sample_alias, levels = use_samples),
    group = factor(group, levels = names(selected_panels))
  )
  

use_colors <- c(
  `Thy1, Syngeneic` = '#F7EC13', 
  `Thy1, NSG`  = '#F7EC13',
  `Puro, Syngeneic` = '#58595B', 
  `Puro, NSG`  = '#58595B',
  `dCas9+Puro, Syngeneic` = 'black',
  `dCas9+Puro, NSG` = 'black'
  )

use_line_colors <- c(
  `Thy1, Syngeneic` = 'black', 
  `Thy1, NSG`  = 'black',
  `Puro, Syngeneic` = 'white', 
  `Puro, NSG`  = 'white',
  `dCas9+Puro, Syngeneic` = 'white',
  `dCas9+Puro, NSG` = 'white'
  )

use_assay %>% 
  ggplot(aes(x=cpm + 1, y = sample_alias, fill = group, color = group, height = after_stat(density))) +
  geom_density_ridges(size = one_pt/4, 
                      alpha = 0.8, 
                      scale = 4,
                      rel_min_height = 0.001 # set the `rel_min_height` argument to remove tails,
                      ) +
  scale_fill_manual(values = use_colors) +
  scale_color_manual(values = use_line_colors) +
  scale_x_log10(expand = expansion(mult = c(0, 0))) +
  scale_y_discrete(expand = c(0.01, 0)) +
  labs(y = '') +
  guides(fill = 'none', color = 'none') +
  facet_wrap2(vars(group), ncol = 2,
                     scales = 'free_y',
                     axes = "x") +
  theme_ridges(font_size = 8, grid = F) +
  theme(
    strip.background = element_blank(),
    axis.line.x = element_line(linewidth = one_pt/4, color = 'black'),
    axis.ticks.x = element_line(linewidth = one_pt/4, color = 'black'),
    axis.text.y = element_blank(),
    strip.text.x = element_text(size=4, hjust = 0)
  )
                                
```


### Supplementary Figure 2B: sgRNA distribution in mice transplanted with Py2T and MVT1 cells
```{r test-dist-tumor-distribution-ridges_plot-py2t-mvt1, fig.width=3, fig.asp=1.33}
selected_panels <- list(
  `Py2T, Thy1, NSG` = c('PY_TumoT_N5', 'PY_TumoT_N4', 'PY_TumoT_N3'),
  `Py2T, Thy1, Syngeneic` = c('PY_TumoT_F3', 'PY_TumoT_F4', 'PY_TumoT_F5'), 
  
  `Py2T, Puro, NSG` = c('PY_TumoP_N1', 'PY_TumoP_N2', 'PY_TumoP_N3'),
  `Py2T, Puro, Syngeneic` = c('PY_TumoP_F1', 'PY_TumoP_F5', 'PY_TumoP_F2'),
  
  `MVT1, Thy1, NSG` = c('MVT_TumoT_N3', 'MVT_TumoT_N4', 'MVT_TumoT_N5'),
  `MVT1, Thy1, Syngeneic` = c('MVT_TumoT_F1', 'MVT_TumoT_F4', 'MVT_TumoT_F3'),
  
  `MVT1, Puro, NSG` = c('MVT_TumoP_N3', 'MVT_TumoP_N4', 'MVT_TumoP_N5'),
  `MVT1, Puro, Syngeneic` = c('MVT_TumoP_F3', 'MVT_TumoP_F4', 'MVT_TumoP_F5')
)

selected_panels_df <- foreach(i = names(selected_panels), .combine = rbind) %do% {
  data.frame(group = i, sample_alias = selected_panels[[i]])
}

use_samples <- unlist(selected_panels)
use_assay <- assay(se[,use_samples], 'cpm') %>% 
  as.data.frame(check.names = F) %>% 
  rownames_to_column('guide') %>% 
  pivot_longer(-guide, names_to = 'sample_alias', values_to = 'cpm') %>% 
  left_join(use_colData) %>% 
  left_join(selected_panels_df) %>% 
  mutate(
    sample_alias = factor(sample_alias, levels = use_samples),
    group = factor(group, levels = names(selected_panels))
  )
  

use_colors <- c(
  `Py2T, Thy1, Syngeneic` = '#F7EC13', 
  `Py2T, Thy1, NSG`  = '#F7EC13',
  `Py2T, Puro, Syngeneic` = '#58595B', 
  `Py2T, Puro, NSG`  = '#58595B',
  
  `MVT1, Thy1, Syngeneic` = '#F7EC13', 
  `MVT1, Thy1, NSG`  = '#F7EC13',
  `MVT1, Puro, Syngeneic` = '#58595B', 
  `MVT1, Puro, NSG`  = '#58595B'
  )

use_line_colors <- c(
  `Py2T, Thy1, Syngeneic` = 'black', 
  `Py2T, Thy1, NSG`  = 'black',
  `Py2T, Puro, Syngeneic` = 'white', 
  `Py2T, Puro, NSG`  = 'white',
  
  `MVT1, Thy1, Syngeneic` = 'black', 
  `MVT1, Thy1, NSG`  = 'black',
  `MVT1, Puro, Syngeneic` = 'white', 
  `MVT1, Puro, NSG`  = 'white'
  )

use_assay %>% 
  ggplot(aes(x=cpm + 1, y = sample_alias, fill = group, color = group, height = after_stat(density))) +
  geom_density_ridges(size = one_pt/4, 
                      alpha = 0.8, 
                      scale = 4,
                      rel_min_height = 0.001 # set the `rel_min_height` argument to remove tails,
                      ) +
  scale_fill_manual(values = use_colors) +
  scale_color_manual(values = use_line_colors) +
  scale_x_log10(expand = expansion(mult = c(0, 0))) +
  scale_y_discrete(expand = c(0.01, 0)) +
  labs(y = '') +
  guides(fill = 'none', color = 'none') +
  facet_wrap2(vars(group), ncol = 2,
                     scales = 'free_y',
                     axes = "x") +
  theme_ridges(font_size = 3, grid = F) +
  theme(
    strip.background = element_blank(),
    axis.line.x = element_line(linewidth = one_pt/4, color = 'black'),
    axis.ticks.x = element_line(linewidth = one_pt/4, color = 'black'),
    axis.text.y = element_blank(),
    panel.border = element_rect(fill =NULL,
                                color = "black",
                                linewidth = 2*one_pt)
  )
                                
```




## Kolmogorov-Smirnov test and ECDF plots 

Run Kolmogorov-Smirnov test to compare distributions merged by group and create Empirical Cumulative Density Function (ECDF) by comparison.
```{r modality-comparisons-ks-ecdf-run}
use_colors <- c(
  `Thy1` = '#F7EC13',
  `Puro`  = '#58595B',
  `dCas9 + Puro` = 'black'
  )

i <- names(comp_list_modality)[1]
ks_test_df <- data.frame()
ecdf_plots <- list()
for(i in names(comp_list_modality)) {
  use_samples <- comp_list_modality[[i]]
  use_assay_c <- assay(se[,use_samples[[1]]], 'cpm') %>% 
    data.frame(check.names = F) %>% 
    rownames_to_column('guide') %>% 
    pivot_longer(-guide, values_to = 'cpm',names_to = 'sample_alias') %>% 
    mutate(group = 'case', condition = names(use_samples)[1])
  use_assay_r <- assay(se[,use_samples[[2]]], 'cpm') %>% 
    data.frame(check.names = F) %>% 
    rownames_to_column('guide') %>% 
    pivot_longer(-guide, values_to = 'cpm',names_to = 'sample_alias') %>% 
    mutate(group = 'referene', condition = names(use_samples)[2])
  use_assay <- rbind(use_assay_c, use_assay_r)
  
  # Perform the Kolmogorov-Smirnov test
  ks.res <- ks.test(use_assay_c$cpm, use_assay_r$cpm)
  ks_test_df <- rbind(
    ks_test_df,
      data.frame(
        comparison = i,
        case = names(use_samples)[1],
        reference = names(use_samples)[2],
        D = ks.res$statistic,
        P = ks.res$p.value
      )
    )

  # Create ECDFs plot Empirical Cumulative Density Function
  # use_colors <- c('red', 'blue') %>% set_names(names(use_samples))
  ecdf_plots[[i]] <- use_assay %>% 
    ggplot(aes(x = log(1+cpm), color = condition, group = sample_alias)) +
    stat_ecdf(geom = 'step', alpha = 1, linewidth = one_pt/4) +
    scale_color_manual(values = use_colors) +
    labs(x = "log(1+cpm)",
         color = '',
         sub = i#,
         # caption =paste("Kolmogorov-Smirnov test p-value:", round(ks.res$p.value, 3), "\n",
                        # "Kolmogorov-Smirnov test D:", round(ks.res$statistic, 3))
         ) +
    theme(
      text = element_text(size=2),
      axis.line = element_line(linewidth = one_pt/4, color = 'black'),
      axis.ticks = element_line(linewidth = one_pt/4, color = 'black'),
      axis.text.y =  element_text(size=3),
      axis.text.x =  element_text(size=3),
      legend.text =  element_text(size=2),
      legend.title =  element_text(size=2),
      plot.subtitle = element_text(size=3),
      plot.caption =  element_text(size=2)
      
    )

}
```

### Supplementary Figure 2C left : K-S test table 
```{r modality-comparisons-ks-table}
ks_test_df %>% 
  dplyr::select(-case, -reference) %>% 
  mutate(P = format.pval(P)) %>% 
  datatable(., 
            rownames = FALSE, 
            filter = 'top', 
            caption = 'Results of Kolmogorov-Smirnov test',
            extensions = 'Buttons', 
            options = list(
              dom = 'Blfrtip',
              buttons = c('csv', 'excel'),
              columnDefs = list(
                list(width = 120, targets = 1)
              )
            )
            ) %>% 
    formatRound(columns = c('D'), digits = 3)
```

### Supplementary Figure 2C right : Empirical Cumulative Density Function (ECDF) plots
Empirical Cumulative Density Function (ECDF) plot by comparison
```{r modality-comparisons-ks-ecdf, fig.width = 3, fig.asp = 0.3}
# Dim for an individual plot : fig.width = 1.5, fig.asp = 0.66
# for(i in names(ecdf_plots)) {
#   cat("#####", i, "\n")
#   print(ecdf_plots[[i]])
#   cat("\n\n")
# }
use_index <- c("Syngeneic Thy1 vs. Puro", "Syngeneic Thy1 vs. dCas9 + Puro")
plot_grid(plotlist = ecdf_plots[use_index], ncol = 2)
```

## Hartigans' dip test
Hartigans' dip test for unimodality. If the p-value is less than 0.05, there's enough evidence to claim that the data are not unimodal (multimodal, at least bimodal).

```{r test-dip}
use_assay <- assay(se, 'cpm')
i <- colnames(use_assay)[1]
modality_test_df <- foreach(i= colnames(use_assay), .combine = rbind) %do%{
  x <- use_assay[,i]
  dip.res <- dip.test(x, simulate.p.value = FALSE, B = 2000)
  data.frame(
    sample_alias = i,
    dip.D = dip.res$statistic,
    dip.p = dip.res$p.value,
    dip.result = ifelse(dip.res$p.value < 0.05, 'multimodal', 'unimodal')
    )
}

modality_test_df %<>% 
  left_join(group_df_modality) %>% 
  left_join(colData(se) %>% data.frame)
  
```

### Supplementary Figure 2D left: Dip test table for Syngeneic models
```{r test-dip-synergeneic-table}
modality_test_df %>% 
  filter(!is.na(modality_group)) %>% 
  filter(mouse_model_global == 'immunocompetent') %>% 
  mutate(
    modality_group = factor(modality_group, levels = names(group_list_modality))
  ) %>%
  group_by(modality_group) %>%
  summarise(
    `Median D` = median(dip.D) %>% round(3),
    n = n(),
    `Multimodal samples (n)` = sum(dip.result == 'multimodal'),
    `Unimodal samples (n)` = sum(dip.result == 'unimodal'),
    `Multimodal samples (%)` = round(100*`Multimodal samples (n)`/n, 1),
    `Unimodal samples (%)` = round(100*`Unimodal samples (n)`/n, 1)
  ) %>% 
  data.frame(check.names = F)
```

### Supplementary Figure 2D right: Dip test table for NSG models
```{r test-dip-nsg-table}
modality_test_df %>% 
  filter(!is.na(modality_group)) %>% 
  filter(mouse_model_global == 'immunodeficient') %>% 
  mutate(
    modality_group = factor(modality_group, levels = names(group_list_modality))
  ) %>%
  group_by(modality_group) %>%
  summarise(
    `Median D` = median(dip.D) %>% round(3),
    n = n(),
    `Multimodal samples (n)` = sum(dip.result == 'multimodal'),
    `Unimodal samples (n)` = sum(dip.result == 'unimodal'),
    `Multimodal samples (%)` = round(100*`Multimodal samples (n)`/n, 1),
    `Unimodal samples (%)` = round(100*`Unimodal samples (n)`/n, 1)
  ) %>% 
  data.frame(check.names = F)
```









## Muller plots

Define selected comparisons for muller plots
```{r define-comparisons-muller-plots}
comp_list_muller_syngeneic_representative <- list(
 `Thy1` = c('4T_TumoT_B3', '4T_LungsT_B3'),
 `Puro` = c('4T_TumoP_B1', '4T_LungsP_B1'),
 `dCas9 + Puro` = c('PlatePositionA3', 'PlatePositionB3')
)
```

Generate muller plots
```{r run-muller-syngeneic-representative-plots}
cpm_threshold <- 10000
comp_list <- comp_list_muller_syngeneic_representative

use_df <- assay(se, 'cpm') %>% data.frame(check.names = FALSE) %>% 
  rownames_to_column('guide') %>% 
  mutate(
    gene = rowData(se)$Gene,
    gene = ifelse(grepl('Non_Target', gene), 'Non_Target', gene)
  ) %>%
  pivot_longer(-c(guide, gene), names_to = 'sample_alias', values_to = 'cpm')

j <-  names(comp_list)[1]
muller_plots <- foreach(j = names(comp_list)) %do% {
  i <- comp_list[[j]]
  x <- use_df %>% 
    filter(sample_alias %in% i) %>% 
    mutate(
      sample_type = ifelse(sample_alias == i[1], 'Primary tumor', 'Lungs'),
      sample_type = factor(sample_type, levels = c('Primary tumor', 'Lungs'))
    )
  
  # Select guides with cpm > cpm_threshold
  x_guides_keep <- x %>% 
    group_by(sample_type, guide) %>% 
    summarise(cpm = max(cpm)) %>% 
    filter(cpm > cpm_threshold) %>% 
    pull(guide)
  
  # Group guides with cpm < cpm_threshold into the low abundant category, and sum counts for low abundant
  x %<>% 
    mutate(
      guide = ifelse(guide %in%  x_guides_keep, guide, 'low abundant')
    ) %>% 
    group_by(sample_type, guide) %>%
    summarise(cpm = sum(cpm)) %>% 
    ungroup() %>%
    mutate(
      guide = fct_reorder(guide, cpm),
      guide = relevel(guide, ref = 'low abundant')
    )
  
  use_cols <- c(
    'low abundant' = 'grey80',
    colorRampPalette(rev(palette_OkabeIto))(nlevels(x$guide) - 1) %>% set_names(levels(x$guide)[-1])
  )
  
  x %>% 
    ggplot( aes(x = sample_type, y = cpm/10000, group = guide, fill = guide)) + 
    geom_area(colour = alpha("white", 0.1), linewidth = 0.08, alpha = 0.8) +
    scale_fill_manual(values = use_cols, guide = guide_legend(ncol = 3)) +
    labs(
      x = '',
      y = expression("CPM x 10"^-4),
      title = j,
      fill = ''
    ) +
    scale_x_discrete(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 1000000/10000)) +
    guides(fill = FALSE) +
    theme(
      panel.background = element_rect(fill = "grey90"),
      # plot.margin = margin(0.5, 1.5, 0.5, 0, "cm"),
      plot.margin = margin(0, 0.5, 0, 0, "cm"),
      axis.line = element_line(linewidth = one_pt/4, color = 'black'),
      axis.ticks = element_line(linewidth = one_pt/4, color = 'black'),
      axis.text =  element_text(size=3),
      axis.title =  element_text(size=3),
      plot.title = element_text(size=3, hjust = 0.5)
    )

}
names(muller_plots) <- names(comp_list)
muller_syngeneic_representative <- muller_plots
```

### Figure 2C: Representative Müller plots of 4T1 in syngeneic recipients

```{r muller-syngeneic-representative-plots, fig.width = 1.5, fig.asp = 2}
wrap_plots(muller_plots, ncol = 1)
```





## Stop knitr
```{r}
knitr::knit_exit()
```

