---
title: "StealTHY CRISPR screen"
subtitle: "Mouse libray (n = 2215) - Reviewer 2 Observation C"
author: "Francesc Castro-Giner"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params:
  data_dir: ./data/crispr/mm_2215_sgRNA
  output_dir: ./output/crispr/mm_2215_sgRNA
  mageck_dir: mageck_rra_r2oC
---

## Load libraries, additional functions and data

Setup environment
```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(results='asis', echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, fig.align = 'center', fig.width = 3.5, fig.asp = 0.618, dpi = 600, dev = c("png", "pdf"), fig.showtext = FALSE, engine.opts = list(bash = "-l"))

options(stringsAsFactors = FALSE)

use_seed <- 1100101

set.seed(use_seed)

# dir.create(params$output_dir, recursive = TRUE, showWarnings = FALSE)
```

Load packages
```{r load-libraries}
library(tidyverse)
library(knitr)
library(foreach)
library(magrittr)
library(DT)
library(kableExtra)
library(diptest)
library(SummarizedExperiment)

library(ggridges)
library(ggh4x)
library(patchwork)
library(colorblindr)
library(ggbeeswarm)
library(ggpubr)

library(MAGeCKFlute)
```

Load ggplot theme
```{r ggplot-theme}
source("./configuration/rmarkdown/ggplot_theme.R")
```

Load ggplot theme
```{r color-palette}
source("./configuration/rmarkdown/color_palettes.R")
```

Load Summarized Experiment object
```{r load-se}
se <- readRDS(file.path(params$data_dir, 'se.rds'))
```

Filter samples used for this analysis
```{r dw-sel-cols}
use_cols <- colData(se) %>% 
  data.frame %>% 
  rownames

se <- se[,use_cols]
```

Data wrangling
```{r dw-colData}
chunk_colData <- colData(se) %>% data.frame %>%
  mutate(
    vector_short = ifelse(vector_short == 'Puro_EF1alpha_SpCas9', 
                          'Cas9+Puro', vector_short),
    vector_short = factor(vector_short, levels = c('Thy1', 'Puro', 'Cas9+Puro')),
    mouse_model_short = case_match(mouse_model_global,
                                   'immunodeficient' ~ 'NSG',
                                   'immunocompetent' ~ 'Syngeneic'
                                   ),
    mouse_model_short = factor(mouse_model_short, levels = c('NSG', 'Syngeneic')),
    vector_mhost = paste0(vector_short, ', ', mouse_model_short),
    model_vector_mhost = paste0(donor,', ', vector_short, ', ', mouse_model_short),
    sample_type = ifelse(cancer_type %in% c('Breast', 'CRC'),
                         case_match(sample_type, 
                             'whole_tumor' ~ 'Primary tumor', 
                             'whole_lung' ~ 'Lung mets',
                             'bulk' ~ 'Bulk',
                             'cell_culture' ~ 'Cell culture'),
                         sample_type
                         ),
        sample_type = ifelse(cancer_type %in% c('NSCLC'),
                         case_match(sample_type, 
                             'met_liver' ~ 'Liver mets', 
                             'whole_lung' ~ 'Primary tumor',
                             'bulk' ~ 'Bulk',
                             'cell_culture' ~ 'Cell culture'),
                         sample_type
                         ),
    
    sample_type = factor(sample_type, levels = c('Primary tumor', 'Lung mets', 'Liver mets', 'Bulk', 'Cell culture'))
    
  )

colData(se) <- chunk_colData %>% DataFrame
```

## Configure analyses

### MAGeCK analysis

#### List of comparisons
List of comparisons for differential abundance using MAGeCK
```{r mageck-comparisons-definition}
x <- colData(se) %>% data.frame
mageck_comp_list <- list(
  `StealTHY KO Syngeneic CMT167` = list(
    `Primary tumor` = x %>% filter(vector_mmodel == 'CMT167-KO-Thy1-C57BL6' & sample_type == 'Primary tumor' ) %>% pull(sample_alias),
    `Liver mets` = x %>% filter(vector_mmodel == 'CMT167-KO-Thy1-C57BL6' & sample_type == 'Liver mets' ) %>% pull(sample_alias)
    ),
  `StealTHY KO Syngeneic LLC1` = list(
    `Primary tumor` = x %>% filter(vector_mmodel == 'LLC1-KO-Thy1-C57BL6' & sample_type == 'Primary tumor' ) %>% pull(sample_alias),
    `Liver mets` = x %>% filter(vector_mmodel == 'LLC1-KO-Thy1-C57BL6' & sample_type == 'Liver mets' ) %>% pull(sample_alias)
    ),
  `StealTHY KO Syngeneic CT26` = list(
    `Primary tumor` = x %>% filter(vector_mmodel == 'CT26-KO-Thy1-BALB' & sample_type == 'Primary tumor' ) %>% pull(sample_alias),
    `Liver mets` = x %>% filter(vector_mmodel == 'CT26-KO-Thy1-BALB' & sample_type == 'Lung mets' ) %>% pull(sample_alias)
    )
)


mageck_comp_list_names <- names(mageck_comp_list) 
names(mageck_comp_list_names) <- names(mageck_comp_list) %>% gsub(" ", "_",.) %>% tolower
```

```{r mageck-comparisons-table}
data_comp <- foreach(i = names(mageck_comp_list), .combine = rbind) %do% {
  c(
    Comparison = i,
    Case = paste(mageck_comp_list[[i]][[1]], collapse = " "),
    Control = paste(mageck_comp_list[[i]][[2]], collapse = " ")
  )
}

data_comp %>% 
  datatable(., 
            rownames = FALSE, 
            filter = 'top', 
            caption = 'List of comparisons for MAGeCK analysis',
            extensions = 'Buttons', 
            options = list(
              dom = 'Blfrtip',
              buttons = c('csv', 'excel'),
              columnDefs = list(
                list(width = 120, targets = 1:2)
              )
            )
            )

```

#### Run MAGeCK

Generate MAGeCK test scripts. To generate the script files change `eval = FALSE` to `eval = TRUE` in the chunk options.

The scripts must be run in the terminal, inside the path `r file.path(params$output_dir, params$ne_mageck_dir)`. The script will activate the mageck environment and run the test for each comparison. The results will be saved in the same directory.
```{r generate-mageck-cmds, eval = FALSE}
chunk_se <- se
out_dir <- file.path(params$output_dir, params$mageck_dir)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# Generate count matrix
count_mat <- assay(chunk_se, 'counts') %>% data.frame(check.names = FALSE) %>% 
  rownames_to_column('sgRNA') %>% 
  mutate(Gene = rowData(chunk_se)$Gene) %>% 
  dplyr::select(sgRNA, Gene, everything())
write_tsv(count_mat, file = file.path(out_dir, 'counts.txt'))

# Generate list of control sgRNA IDs
non_target_list <- rownames(chunk_se) %>% grep('Non_Target', ., value = T) %>% data.frame
write_tsv(non_target_list, file = file.path(out_dir, 'control_sgrna.txt'), col_names = F)

# Generate mageck commands
res_cmd <- data.frame('source activate mageckenv')
res_cmd <- foreach(i=names(mageck_comp_list), .combine = rbind) %do% {
  paste('mageck test -k counts.txt --control-sgrna control_sgrna.txt --norm-method control -t', paste(mageck_comp_list[[i]][[1]], collapse = ','),'-c',  paste(mageck_comp_list[[i]][[2]], collapse = ','), '-n', tolower(gsub(" ", "_", i)), '\n')
} %>% data.frame()

res_cmd <- rbind('conda activate mageckenv', res_cmd, 'conda deactivate')
write_tsv(res_cmd, file.path(out_dir, 'run_mageck_test.sh'), col_names = FALSE)
```

#### Load MAGeCK results
```{r load-mageck}
gene_summ_files <- list.files(path = file.path(params$output_dir, params$mageck_dir), pattern = 'gene_summary.txt', full.names = TRUE)
analysis_prefix <- basename(gene_summ_files) %>% gsub(".gene_summary.txt", "", .)
gene_summ <- foreach(i = gene_summ_files) %do% read.delim(i, check.names = FALSE)
names(gene_summ) <- analysis_prefix

sgrna_summ_files <- list.files(path = file.path(params$output_dir, params$mageck_dir), pattern = 'sgrna_summary.txt', full.names = TRUE)
analysis_prefix <- basename(sgrna_summ_files) %>% gsub(".sgrna_summary.txt", "", .)
sgrna_summ <- foreach(i = sgrna_summ_files) %do% read.delim(i, check.names = FALSE)
names(sgrna_summ) <- analysis_prefix

gene_summ_files <- list.files(path = file.path(params$output_dir, params$mageck_dir), pattern = 'gene_summary.txt', full.names = TRUE)
analysis_prefix <- basename(gene_summ_files) %>% gsub(".gene_summary.txt", "", .)
gene_summ <- foreach(i = gene_summ_files) %do% read.delim(i, check.names = FALSE)
names(gene_summ) <- analysis_prefix


gene_summ_files <- list.files(path = file.path('./output/crispr/mm_2215_sgRNA', 'mageck_rra_expr'), pattern = 'gene_summary.txt', full.names = TRUE)
analysis_prefix <- basename(gene_summ_files) %>% gsub(".gene_summary.txt", "", .)
gene_summ_5C <- foreach(i = gene_summ_files) %do% read.delim(i, check.names = FALSE)
names(gene_summ_5C) <- analysis_prefix
```


## Differential abundance analysis
```{r da-selected-genes-conf}
gene_info <- readxl::read_xlsx('./data/resources/selected_gene_list/gene_list_for_figure_5c.xlsx')

use_colors <- c(
  `Immune system cytokines and chemokines` = 'grey',
  `Immune system cytokines\nand chemokines` = 'grey',
  
  `Interferone, TNF and danger/death-inducing signals` = 'black',
  `Interferon, TNF\nand danger/death-inducing\nsignals` = 'black',
  `Interferon, TNF and\ndanger/death-inducing\nsignals` = 'black',
  
  `TGF-β family related` = 'red',
  
  `RTK signaling` = 'blue',
  `Positive controls` = 'black'
)

use_fill <- c(
  `Immune system cytokines and chemokines` = 'grey',
  `Immune system cytokines\nand chemokines` = 'grey',
  
  `Interferone, TNF and danger/death-inducing signals` = 'black',
  `Interferon, TNF\nand danger/death-inducing\nsignals` = 'black',
  `Interferon, TNF and\ndanger/death-inducing\nsignals` = 'black',
  `TGF-β family related` = 'red',
  
  `RTK signaling` = 'blue',
  `Positive controls` = 'white'
)

use_shape <- c(
  `Immune system cytokines and chemokines` = 16,
  `Immune system cytokines\nand chemokines` = 16,
  
  `Interferone, TNF and danger/death-inducing signals` = 16,
  `Interferon, TNF\nand danger/death-inducing\nsignals` = 16,
  `Interferon, TNF and\ndanger/death-inducing\nsignals` = 16,
  `TGF-β family related` = 16,
  
  `RTK signaling` = 16,
  `Positive controls` = 1
)

# get gene order based on Original figure
# gene_levels <- gene_summ_5C[['stealthy_ko_syngeneic']] %>%
#   ReadRRA(score = 'lfc') %>%
#   dplyr::rename(LFC = Score) %>%
#   left_join(gene_info, by = c('id' = 'id')) %>%
#   filter(!is.na(gene_set)) %>%
#   mutate(gene_name = fct_reorder(gene_name, LFC, .desc = FALSE)) %>%
#   pull(gene_name) %>%
#   levels
gene_levels <- rev(gene_info$id)
```

### Differential abundance of selected genes
Differential abundance of selected genes in carcinoma cells subjected to StealTHY KO or Cas9+Puro

```{r da-selected-genes-dotplot, fig.width=2.37, fig.asp=2.2}
use_comp <- names(gene_summ)

# FDR limits
xfdr <- foreach(i=gene_summ[use_comp], .combine = rbind) %do% {
  i %>% 
    ReadRRA(score = 'lfc') %>% 
    dplyr::rename(LFC = Score) %>% 
    left_join(gene_info, by = c('id' = 'id')) %>% 
    filter(!is.na(gene_set)) %>% 
    pull(FDR) %>% 
    min
}
maxFDR <- -log10(xfdr[,1]) %>% max
 
# Set LFC (x-axis) for comparable plots to "stealthy_ko_syngeneic": "stealthy_ko_nsg"
# stealthy_comp <- c("stealthy_ko_syngeneic", "stealthy_ko_nsg")
# xlfc <- foreach(i=gene_summ_5C[stealthy_comp], .combine = rbind) %do% {
#   i %>% 
#     ReadRRA(score = 'lfc') %>% 
#     dplyr::rename(LFC = Score) %>% 
#     left_join(gene_info, by = c('id' = 'id')) %>% 
#     filter(!is.na(gene_set)) %>% 
#     pull(LFC) %>% 
#     abs %>% max
# }
# stealthy_maxLFC <- max(xlfc[,1])
# stealthy_xlim <- c(-stealthy_maxLFC, stealthy_maxLFC)
# stealthy_breaks <- c(-2.5, -1, 0, 1, 2.5)
# stealthy_labels <- c('-2.5', '-1', '0', '1', '2.5')

xlfc <- foreach(i=gene_summ, .combine = rbind) %do% {
  i %>%
    ReadRRA(score = 'lfc') %>%
    dplyr::rename(LFC = Score) %>%
    left_join(gene_info, by = c('id' = 'id')) %>%
    filter(!is.na(gene_set)) %>%
    pull(LFC) %>%
    abs %>% max
}
maxLFC <- ceiling(max(xlfc[,1]))
xlim <- c(-maxLFC, maxLFC)
xbreaks <- c(-6, -2.5, 0, 2.5, 6)
xlabels <- c('-6', '-2.5', '0', '2.5', '6')

i <- use_comp[1]
for(i in use_comp)  {
  cat("####", mageck_comp_list_names[i], "\n\n")
  
  x <- gene_summ[[i]] %>% 
    ReadRRA(score = 'lfc') %>% 
    dplyr::rename(LFC = Score) %>% 
    left_join(gene_info, by = c('id' = 'id')) %>% 
    filter(!is.na(gene_set)) %>% 
    mutate(
      gene_name = factor(gene_name, levels = gene_levels),
      gene_set = ifelse(
        gene_set == 'Immune system cytokines and chemokines',
        'Immune system cytokines\nand chemokines',
        gene_set
      ),
      gene_set = ifelse(
        gene_set == 'Interferone, TNF and danger/death-inducing signals',
        'Interferon, TNF and\ndanger/death-inducing\nsignals',
        gene_set
      ),
      gene_set = factor(gene_set, names(use_colors))
      )
  # xlim <- c(-ceiling(max(abs(x$LFC), na.rm = TRUE)), 
  #           ceiling(max(abs(x$LFC), na.rm = TRUE)))
  # size_breaks <- c(0, 0.25, 0.5, 1, 2, ceiling(maxFDR))
  size_breaks <- c(0, 0.25, 0.5, ceiling(maxFDR))
  res <- x %>% 
    ggplot(aes(LFC, gene_name, color = gene_set, size = -log10(FDR))) +
    geom_point(aes(shape = gene_set)) +
    geom_vline(xintercept = 0, linetype = 3, linewidth = one_pt/2) +
    scale_y_discrete(expand = expansion(add = c(1, 1))) + # add fixed units around each facet to avoid lines over points
    # scale_x_continuous(
    #   expand = expansion(add = c(1, 1))
    #   ) + # add fixed units around each facet to avoid lines over points
    scale_color_manual(values = use_colors) +
    scale_size(range = c(0.5, 2.5), 
               limits = c(0, ceiling(maxFDR)), 
               breaks = size_breaks,
               labels = size_breaks %>% as.character # remove trailing 0 from decimals
               ) +
    scale_shape_manual(values = use_shape) +
    scale_x_continuous(
        expand = expansion(add = c(1, 1))
        ) +
    facet_grid(rows = vars(gene_set), scales = 'free_y', space = 'free_y') +
    theme_facet +
    theme(
      # axis.line = element_line(linewidth = one_pt/2, color = 'black'),
      axis.ticks = element_line(linewidth = one_pt/2, color = 'black'),
      panel.border = element_rect(linewidth = one_pt/2, color = 'black'),
      axis.text = element_text(size=8), 
      axis.title = element_text(size = 8),
      legend.title = element_text(size=8),
      legend.text = element_text(size=8),
      strip.text = element_text(size = 4)
    ) +
    labs(
      x = expression(paste("lo", g[2],"(Fold change)")),
      y = NULL,
      size = expression(paste("-lo", g[10],"(FDR)")),
      title = 
    ) +
    guides(color = 'none', shape = 'none')
  
  res <- res +
    scale_x_continuous(
      expand = expansion(add = c(1, 1)), 
      limits = xlim,
      breaks = xbreaks,
      labels = xlabels
    )
  
  print(res)
  cat('\n\n')
}
```


## Muller plots of selected genes

Define selected comparisons for muller plots
```{r define-comparisons-muller-plots}

mageck_comp_list <- list(
  `StealTHY KO Syngeneic CMT167` = list(
    `Primary tumor` = x %>% filter(vector_mmodel == 'CMT167-KO-Thy1-C57BL6' & sample_type == 'Primary tumor' ) %>% pull(sample_alias),
    `Liver mets` = x %>% filter(vector_mmodel == 'CMT167-KO-Thy1-C57BL6' & sample_type == 'Liver mets' ) %>% pull(sample_alias)
    ),
  `StealTHY KO Syngeneic LLC1` = list(
    `Primary tumor` = x %>% filter(vector_mmodel == 'LLC1-KO-Thy1-C57BL6' & sample_type == 'Primary tumor' ) %>% pull(sample_alias),
    `Liver mets` = x %>% filter(vector_mmodel == 'LLC1-KO-Thy1-C57BL6' & sample_type == 'Liver mets' ) %>% pull(sample_alias)
    ),
  `StealTHY KO Syngeneic CT26` = list(
    `Primary tumor` = x %>% filter(vector_mmodel == 'CT26-KO-Thy1-BALB' & sample_type == 'Primary tumor' ) %>% pull(sample_alias),
    `Liver mets` = x %>% filter(vector_mmodel == 'CT26-KO-Thy1-BALB' & sample_type == 'Lung mets' ) %>% pull(sample_alias)
    )
)


comp_list_muller <- list(
  `Comparison 1` = list(
    `CMT167 Syngeneic Thy1.1 1` = c('C167_StealTHY_Tum_1',	'C167_StealTHY_Met_1'),
    `CMT167 Syngeneic Thy1.1 3` = c('C167_StealTHY_Tum_3',	'C167_StealTHY_Met_3'),
    `CMT167 Syngeneic Thy1.1 4` = c('C167_StealTHY_Tum_4',	'C167_StealTHY_Met_4'),
    `CMT167 Syngeneic Thy1.1 5` = c('C167_StealTHY_Tum_5',	'C167_StealTHY_Met_5')
  ),
  `Comparison 2` = list( 
   `LLC1 Syngeneic Thy1.1 1` =  c('LLC1_StealTHY_Tum_1',	'LLC1_StealTHY_Met_1'),
   `LLC1 Syngeneic Thy1.1 2` =  c('LLC1_StealTHY_Tum_2',	'LLC1_StealTHY_Met_2'),
   `LLC1 Syngeneic Thy1.1 3` =  c('LLC1_StealTHY_Tum_3',	'LLC1_StealTHY_Met_3'),
   `LLC1 Syngeneic Thy1.1 4` =  c('LLC1_StealTHY_Tum_4',	'LLC1_StealTHY_Met_4'),
   `LLC1 Syngeneic Thy1.1 5` =  c('LLC1_StealTHY_Tum_5',	'LLC1_StealTHY_Met_5')
  ),
  `Comparison 3` = list( 
    `CT26 Syngeneic Thy1.1 1` = c('ColoCT26_StealTHY_Tum_1',	'ColoCT26_StealTHY_Met_1'),
    `CT26 Syngeneic Thy1.1 2` = c('ColoCT26_StealTHY_Tum_2',	'ColoCT26_StealTHY_Met_2'),
    `CT26 Syngeneic Thy1.1 3` = c('ColoCT26_StealTHY_Tum_3',	'Replacement_1'),
    `CT26 Syngeneic Thy1.1 4` = c('Replacement_2',	'Replacement_3'),
    `CT26 Syngeneic Thy1.1 5` = c('Replacement_4',	'Replacement_4bis')
  )
)
genes_list_1 <- gene_info %>% 
  filter(gene_set %in% c("TGF-β family related", "RTK signaling", "Positive controls")) %>% 
  pull(id)
genes_list_2 <- gene_info %>% 
  filter(!gene_set %in% c("TGF-β family related", "RTK signaling", "Positive controls")) %>% 
  pull(id)

```

Generate muller plots for genes in main figure 5C
```{r run-muller-plots}
cpm_threshold <- 10000
comp_list <- comp_list_muller
use_genes <- genes_list_1

use_df <- assay(se, 'cpm') %>% data.frame(check.names = FALSE) %>% 
  rownames_to_column('guide') %>% 
  mutate(
    gene = rowData(se)$Gene,
    gene = ifelse(grepl('Non_Target', gene), 'Non_Target', gene)
  ) %>%
  filter(gene %in% use_genes) %>% 
  pivot_longer(-c(guide, gene), names_to = 'sample_alias', values_to = 'cpm')

j <-  names(comp_list)[1]
muller_plots <- foreach(j = names(comp_list)) %do% {
  i <- comp_list[[j]][[1]]
  iname <-  names(comp_list[[j]])[1]
  res <- foreach(iname = names(comp_list[[j]])) %do% {
    i <- comp_list[[j]][[iname]]
    x <- use_df %>% 
      filter(sample_alias %in% i) %>% 
      mutate(
        sample_type = ifelse(sample_alias == i[1], 'Primary tumor', 'Metastasis'),
        sample_type = factor(sample_type, levels = c('Primary tumor', 'Metastasis'))
      )
    
    # Group CPM by gene
    x %<>% 
      group_by(gene, sample_alias, sample_type) %>% 
      summarise(cpm = sum(cpm)) %>% 
      mutate(
        gene = fct_reorder(gene, cpm)
      )
    # 
    # # Select guides with cpm > cpm_threshold
    # x_guides_keep <- x %>% 
    #   group_by(sample_type, guide) %>% 
    #   summarise(cpm = max(cpm)) %>% 
    #   filter(cpm > cpm_threshold) %>% 
    #   pull(guide)
    
    # # Group guides with cpm < cpm_threshold into the low abundant category, and sum counts for low abundant
    # x %<>% 
    #   mutate(
    #     guide = ifelse(guide %in%  x_guides_keep, guide, 'low abundant')
    #   ) %>% 
    #   group_by(sample_type, guide) %>%
    #   summarise(cpm = sum(cpm)) %>% 
    #   ungroup() %>%
    #   mutate(
    #     guide = fct_reorder(guide, cpm),
    #     guide = relevel(guide, ref = 'low abundant')
    #   )
    
    use_cols <- c(
      'low abundant' = 'grey80',
      colorRampPalette(rev(palette_OkabeIto))(nlevels(x$gene) - 1) %>% set_names(levels(x$gene)[-1])
    )
    
    x %>% 
      ggplot( aes(x = sample_type, y = cpm/10000, group = gene, fill = gene)) + 
      geom_area(colour = alpha("white", 0.1), linewidth = 0.08, alpha = 0.8) +
      scale_fill_manual(values = use_cols, guide = guide_legend(ncol = 3)) +
      labs(
        x = '',
        y = expression("CPM x 10"^-4),
        title = iname,
        fill = ''
      ) +
      scale_x_discrete(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0), limits = c(0, 1000000/10000)) +
      # guides(fill = 'none') +
      theme(
        panel.background = element_rect(fill = "grey90"),
        # plot.margin = margin(0.5, 1.5, 0.5, 0, "cm"),
        plot.margin = margin(0, 0.5, 0, 0, "cm"),
        axis.text =  element_text(size=3),
        axis.title =  element_text(size=3),
        plot.title = element_text(size=3, hjust = 0.5)
      )
  }
  names(res) <- names(comp_list[[j]])
  return(res)

}
names(muller_plots) <- names(comp_list)
# muller_syngeneic_representative <- muller_plots
```


```{r muller-plots, fig.width = 1.5, fig.asp = 2}
for (i in names(muller_plots)) {
  cat('###', i, '\n\n')
  wrap_plots(muller_plots[[i]], ncol = 1) %>% print
  cat('\n\n')
}

```
