---
title: "Clonal dynamics in humanized NSG mice (huNSG)"
subtitle: "Human libray (n = 2180)"
author: "Francesc Castro-Giner"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
params:
  data_dir: ./data/crispr/hsa_2180_sgRNA
  output_dir: ./output/crispr/hsa_2180_sgRNA
  mageck_dir: mageck_rra
---

## Load libraries, additional functions and data

Setup environment
```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(results='asis', echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, fig.align = 'center', fig.width = 3.5, fig.asp = 0.618, dpi = 600, dev = c("png", "pdf"), fig.showtext = FALSE, engine.opts = list(bash = "-l"))

options(stringsAsFactors = FALSE)

use_seed <- 1100101

set.seed(use_seed)

# dir.create(params$output_dir, recursive = TRUE, showWarnings = FALSE)
```

Load packages
```{r load-libraries}
library(tidyverse)
library(knitr)
library(foreach)
library(magrittr)
library(DT)
library(kableExtra)
library(diptest)
library(SummarizedExperiment)

library(ggridges)
library(ggh4x)
library(patchwork)
library(colorblindr)
library(ggbeeswarm)
library(ggpubr)
library(cowplot)
library(grid)
library(ggrepel)
library(MAGeCKFlute)
```

Load ggplot theme
```{r ggplot-theme}
source("./configuration/rmarkdown/ggplot_theme.R")
```

Load ggplot theme
```{r color-palette}
source("./configuration/rmarkdown/color_palettes.R")
```

Load custom functions
```{r load-functions}
source('./code/R-functions/geom_split_violin.R')
```

Load Summarized Experiment object
```{r load-se}
se <- readRDS(file.path(params$data_dir, 'se.rds'))
```

Filter samples used for this analysis
```{r dw-sel-cols, eval = FALSE}
use_cols <- colData(se) %>% 
  data.frame %>% 
  filter(effect_type_short == 'noKO') %>% 
  filter(vector_short != "Thy1_EF1alpha_dCas9") %>% 
  rownames

se <- se[,use_cols]
```

Data wrangling
```{r dw-colData}
chunk_colData <- colData(se) %>% data.frame %>%
  mutate(
    immunogen_mmmodel = paste(immunogen, mm_model_type, sep = '_'),
    immunogen_burden = ifelse(
      immunogen == 'no_immunogen' & mm_model_type == 'Imm.competent',
      'HuTHY1', NA),
    immunogen_burden = ifelse(
      immunogen == 'dCas9-GFP' & mm_model_type == 'Imm.deficient',
      'GFP', immunogen_burden),
    immunogen_burden = ifelse(
      immunogen == 'dCas9-THY' & mm_model_type == 'Imm.competent',
      'Cas9-HuTHY1', immunogen_burden),
    immunogen_burden = ifelse(
      immunogen == 'dCas9-GFP' & mm_model_type == 'Imm.competent',
      'Cas9-GFP', immunogen_burden),
    immunogen_burden = factor(
      immunogen_burden,
      levels = c('HuTHY1', 'GFP', 'Cas9-HuTHY1', 'Cas9-GFP')),
    sample_type = case_match(
        sample_type,
        "In vitro" ~ 'Input',
        "whole_tumor" ~ 'Primary tumor',
        "whole_lung" ~ 'Lungs',
        "met_lymph" ~ 'Lymph nodes',
        "met_liver" ~ 'Liver'
      ),
    sample_type = factor(
      sample_type,
      levels = c('Input', 'Primary tumor', 'Lungs', 'Liver', 'Lymph nodes')               
    ),
    
    mouse_model_short = ifelse(mm_model_type == 'Imm.competent',
                               'huNSG', NA),
    mouse_model_short = ifelse(mm_model_type == 'Imm.deficient',
                               'NSG', mouse_model_short),
    mouse_model_short = ifelse(mm_model_type == 'In_vitro',
                               'In vitro', mouse_model_short)
    # 
    # vector_short = ifelse(vector_short == 'Puro_EF1alpha_dCas9', 
    #                       'dCas9+Puro', vector_short),
    # vector_short = factor(vector_short, levels = c('Thy1', 'Puro', 'dCas9+Puro')),
    # mouse_model_short = case_match(mouse_model_global,
    #                                'immunodeficient' ~ 'NSG',
    #                                'immunocompetent' ~ 'Syngeneic'
    #                                ),
    # mouse_model_short = factor(mouse_model_short, levels = c('NSG', 'Syngeneic')),
    # vector_mhost = paste0(vector_short, ', ', mouse_model_short),
    # model_vector_mhost = paste0(donor,', ', vector_short, ', ', mouse_model_short),
    # 
    # sample_type = case_match(sample_type, 
    #                          'whole_tumor' ~ 'Primary tumor', 
    #                          'whole_lung' ~ 'Lung mets'),
    # sample_type = factor(sample_type, levels = c('Primary tumor', 'Lung mets'))
  )

colData(se) <- chunk_colData %>% DataFrame

```

## Configure analysis

### MAGeCK analysis

#### List of comparisons
List of comparisons for differential abundance using MAGeCK
```{r mageck-comparisons-definition}
x <- colData(se) %>% data.frame

mageck_comp_list <- list(
  `StealTHY huNSG` = list(
    `Primary tumor` = x %>% filter(immunogen == 'StealTHY' & mm_model_type == 'Imm.competent' & sample_type == 'Primary tumor') %>% 
      arrange(replicate) %>% 
      pull(sample_alias),
    `Lung mets` = x %>% filter(immunogen == 'StealTHY' & mm_model_type == 'Imm.competent' & sample_type == 'Lungs') %>% 
      filter(replicate %in% 1:7) %>% 
      arrange(replicate) %>% 
      pull(sample_alias),
    paired = TRUE
    ),
  `StealTHY NSG` = list(
    Tumor = x %>% filter(immunogen == 'StealTHY' & mm_model_type == 'Imm.deficient' & sample_type == 'Primary tumor') %>% 
      arrange(replicate) %>% 
      pull(sample_alias),
    Lung = x %>% filter(immunogen == 'StealTHY' & mm_model_type == 'Imm.deficient' & sample_type == 'Lungs') %>% 
      arrange(replicate) %>% 
      pull(sample_alias),
    paired = TRUE
    )
)


mageck_comp_list_names <- names(mageck_comp_list) 
names(mageck_comp_list_names) <- names(mageck_comp_list) %>% gsub(" ", "_",.) %>% tolower
```


#### Run MAGeCK

Generate MAGeCK test scripts. To generate the script files change `eval = FALSE` to `eval = TRUE` in the chunk options.

The scripts must be run in the terminal, inside the path `r file.path(params$output_dir, params$mageck_dir)`. The script will activate the mageck environment and run the test for each comparison. The results will be saved in the same directory.
```{r generate-mageck-cmds, eval = FALSE}
out_dir <- file.path(params$output_dir, params$mageck_dir)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# Generate count matrix
count_mat <- assay(se, 'counts') %>% data.frame(check.names = FALSE) %>% 
  rownames_to_column('sgRNA') %>% 
  mutate(Gene = rowData(se)$Gene) %>% 
  dplyr::select(sgRNA, Gene, everything())
write_tsv(count_mat, file = file.path(out_dir, 'counts.txt'))

# Generate list of control sgRNA IDs
non_target_list <- rownames(se) %>% grep('Non_Target', ., value = T) %>% data.frame
write_tsv(non_target_list, file = file.path(out_dir, 'control_sgrna.txt'), col_names = F)

# Generate mageck commands
res_cmd <- data.frame('source activate mageckenv')
i=names(mageck_comp_list)[1]
res_cmd <- foreach(i=names(mageck_comp_list), .combine = rbind) %do% {
  if(mageck_comp_list[[i]]$paired) {
    paste('mageck test -k counts.txt --control-sgrna control_sgrna.txt --norm-method control --paired -t', paste(mageck_comp_list[[i]][[2]], collapse = ','),'-c',  paste(mageck_comp_list[[i]][[1]], collapse = ','), '-n', tolower(gsub(" ", "_", i)), '\n')} else {
    paste('mageck test -k counts.txt --control-sgrna control_sgrna.txt --norm-method control -t', paste(mageck_comp_list[[i]][[2]], collapse = ','),'-c',  paste(mageck_comp_list[[i]][[1]], collapse = ','), '-n', tolower(gsub(" ", "_", i)), '\n')
  }
} %>% data.frame()

res_cmd <- rbind('conda activate mageckenv', res_cmd, 'conda deactivate')
write_tsv(res_cmd, file.path(out_dir, 'run_mageck_test.sh'), col_names = FALSE)

```

#### Load MAGeCK results
```{r load-mageck}
gene_summ_files <- list.files(path = file.path(params$output_dir, params$mageck_dir), pattern = 'gene_summary.txt', full.names = TRUE)
analysis_prefix <- basename(gene_summ_files) %>% gsub(".gene_summary.txt", "", .)
gene_summ <- foreach(i = gene_summ_files) %do% read.delim(i, check.names = FALSE)
names(gene_summ) <- analysis_prefix

sgrna_summ_files <- list.files(path = file.path(params$output_dir, params$mageck_dir), pattern = 'sgrna_summary.txt', full.names = TRUE)
analysis_prefix <- basename(sgrna_summ_files) %>% gsub(".sgrna_summary.txt", "", .)
sgrna_summ <- foreach(i = sgrna_summ_files) %do% read.delim(i, check.names = FALSE)
names(sgrna_summ) <- analysis_prefix
```




## Muller plots

Define muller plots groups
```{r configure-muller-plots}
x <- colData(se) %>% data.frame() 

muller_groups_selected <- list(
  `HuTHY1` = x %>% 
    filter(immunogen_burden == 'HuTHY1' & replicate == '4') %>% 
    pull(sample_alias),
  `GFP` = x %>% 
    filter(immunogen_burden == 'GFP' & replicate == '4') %>% 
    pull(sample_alias),
  `Cas9-HuTHY1` = x %>% 
    filter(immunogen_burden == 'Cas9-HuTHY1' & replicate == '1') %>% 
    pull(sample_alias),
  `Cas9-GFP` = x %>% 
    filter(immunogen_burden == 'Cas9-GFP' & replicate == '1') %>% 
    pull(sample_alias)
)


# Add loading control
loading_control <- 'Replacement_Sample'
muller_groups_selected <- map(muller_groups_selected, function(x) c(loading_control, x))
```

Generate muller plots
```{r run-muller-plots}
cpm_threshold <- 50000

use_df <- assay(se, 'cpm') %>% data.frame(check.names = FALSE) %>% 
  rownames_to_column('guide') %>% 
  mutate(
    gene = rowData(se)$Gene,
    gene = ifelse(grepl('Non_Target', gene), 'Non_Target', gene)
  ) %>%
  pivot_longer(-c(guide, gene), names_to = 'sample_alias', values_to = 'cpm') %>% 
  left_join(colData(se) %>% data.frame %>% dplyr::select(sample_alias, sample_type)) %>% 
  mutate(
    sample_type = recode_factor(
        sample_type,
        `Primary tumor` = 'Primary\ntumor',
        `Lymph nodes` = 'Lymph\nnodes'
      ),
    sample_type = factor(
      as.character(sample_type),
      levels = c('Input', 'Primary\ntumor', 'Lungs', 'Lymph\nnodes', 'Liver')
    )
  )

i <- names(muller_groups_selected)[1]
muller_plots <- foreach(i = names(muller_groups_selected)) %do% {
  j <-  muller_groups_selected[[i]]
  
  x <- use_df %>% 
    filter(sample_alias %in% j) %>% 
    mutate(
      sample_alias = factor(sample_alias, j)
    )
  x_guides_keep <- x %>% group_by(sample_alias, sample_type, guide) %>% summarise(cpm = max(cpm)) %>% filter(cpm > cpm_threshold) %>% pull(guide)
  
  low_abundant_tag <- paste('<', cpm_threshold, 'cpm')
  x %<>% 
    mutate(
      guide = ifelse(guide %in%  x_guides_keep, guide, low_abundant_tag)
    ) %>% 
    group_by(sample_alias, sample_type, guide) %>%
    summarise(cpm = sum(cpm)) %>% 
    ungroup() %>%
    mutate(
      guide = fct_reorder(guide, cpm),
      guide = relevel(guide, ref = low_abundant_tag)
    )
  
  use_cols <- c(
    'grey80',
    colorRampPalette(rev(palette_OkabeIto[1:7]))(nlevels(x$guide) - 1) %>% set_names(levels(x$guide)[-1])
  )
  names(use_cols)[1] <- low_abundant_tag
  
  x %>% 
    ggplot( aes(x = sample_type, y = cpm/10000, group = guide, fill = guide)) + 
    geom_area(colour = alpha("white", 0.1), linewidth = 0.08, alpha = 0.8) +
    scale_fill_manual(values = use_cols, guide = guide_legend(ncol = 3)) +
    labs(
      x = '',
      y = expression("CPM x 10"^-4),
      fill = '',
      title = i
    ) +
    theme(
      # panel.background = element_rect(fill = "grey90"),
      # plot.margin = margin(0.5, 1.5, 0.5, 0, "cm"),
      plot.margin = margin(0, 0.5, 0, 0, "cm"),
      axis.text =  element_text(size=6),
      plot.title = element_text(size=6, hjust = 0.5)
    ) +
    guides(fill = 'none') +
    scale_x_discrete(expand = c(0, 0), guide = guide_axis(angle = 45)) +
    scale_y_continuous(expand = c(0, 0))
}
names(muller_plots) <- names(muller_groups_selected)
```

### Figure 5H: Muller plots for huNSGs
Muller plots showing the relative frequency (CPM) of unique sgRNAs in the indicated samples. Guides with CPM ≥ `r cpm_threshold` are colored. Individual guides with CPM < `r cpm_threshold` are grouped in a grey area.
```{r muller-plots, fig.width = 7.55, fig.asp = 0.19}
wrap_plots(muller_plots, nrow = 1)
```


## sgRNA distribution by condition Primary tumor
```{r test-dist-tumor-conf}
use_sample_type <- 'Primary tumor'
use_colData <- colData(se) %>% data.frame %>%
  filter(sample_type == use_sample_type)
```

### Figure 6B: sgRNA distribution of in mice transplanted with cells subjected to StealTHY KO

```{r test-dist-tumor-distribution-ridges_plot--ic-id, fig.asp = 1.03, fig.width = 1.79}
sample_ord <- rev(c(
  'C1Position_HumanPlate2',
  'C2Position_HumanPlate2',
  'E2Position_HumanPlate2',
  'E1Position_HumanPlate2',
  'G1Position_HumanPlate2',
  
  'C4Position_HumanPlate2',
  'C3Position_HumanPlate2',
  'G3Position_HumanPlate2',
  'E3Position_HumanPlate2',
  'A4Position_HumanPlate2'
  ))

use_assay <- assay(se[,sample_ord], 'cpm') %>% 
  as.data.frame(check.names = F) %>% 
  rownames_to_column('guide') %>% 
  pivot_longer(-guide, names_to = 'sample_alias', values_to = 'cpm') %>% 
  left_join(use_colData) %>% 
  mutate(
    sample_alias = factor(sample_alias, levels = sample_ord)
    )


use_assay %>% 
  ggplot(aes(x=cpm + 1, y = sample_alias, 
             fill = mouse_model_short, 
             height = after_stat(density))) +
  geom_density_ridges(color = 'black',
                      size = one_pt/4, 
                      alpha = 0.8, 
                      scale = 4,
                      rel_min_height = 0.001 # set the `rel_min_height` argument to remove tails,
                      ) +
  scale_fill_manual(values = palette_hu_stealthy_ko) +
  scale_x_log10(expand = expansion(mult = c(0, 0))) +
  scale_y_discrete(expand = c(0.01, 0)) +
  labs(y = '') +
  guides(fill = 'none', color = 'none') +
  theme_ridges(font_size = 6, grid = F) +
  theme(
    strip.background = element_blank(),
    axis.line.x = element_line(linewidth = one_pt/4, color = 'black'),
    axis.ticks.x = element_line(linewidth = one_pt/4, color = 'black'),
    axis.text.y = element_blank()
  )
```


## Differential abundance analysis
```{r diff-abundance-plots-conf}
gene_info <- readxl::read_xlsx('./data/resources/selected_gene_list/gene_list_for_figure_7d.xlsx')

use_colors <- c(
  `Immune system cytokines and chemokines` = 'grey',
  `Immune system cytokines\nand chemokines` = 'grey',
  
  `Interferone, TNF and danger/death-inducing signals` = 'black',
  `Interferon, TNF\nand danger/death-inducing\nsignals` = 'black',
  `Interferon, TNF and\ndanger/death-inducing\nsignals` = 'black',
  `TGF-β family related` = 'red',
  
  `RTK signaling` = 'blue',
  `Positive controls` = 'black'
)

use_fill <- c(
  `Immune system cytokines and chemokines` = 'grey',
  `Immune system cytokines\nand chemokines` = 'grey',
  
  `Interferone, TNF and danger/death-inducing signals` = 'black',
  `Interferon, TNF\nand danger/death-inducing\nsignals` = 'black',
  `Interferon, TNF and\ndanger/death-inducing\nsignals` = 'black',
  `TGF-β family related` = 'red',
  
  `RTK signaling` = 'blue',
  `Positive controls` = 'white'
)

use_shape <- c(
  `Immune system cytokines and chemokines` = 16,
  `Immune system cytokines\nand chemokines` = 16,
  
  `Interferone, TNF and danger/death-inducing signals` = 16,
  `Interferon, TNF\nand danger/death-inducing\nsignals` = 16,
  `Interferon, TNF and\ndanger/death-inducing\nsignals` = 16,
  `TGF-β family related` = 16,
  
  `RTK signaling` = 16,
  `Positive controls` = 1
)

# get gene order based on LFC in StealTHY_IC
i <- 'stealthy_hunsg'
gene_levels <- gene_summ[[i]] %>% 
  ReadRRA(score = 'lfc') %>% 
  dplyr::rename(LFC = Score) %>% 
  left_join(gene_info, by = c('id' = 'id')) %>% 
  filter(!is.na(gene_set)) %>% 
  mutate(gene_name = fct_reorder(gene_name, LFC)) %>% 
  pull(gene_name) %>% 
  levels
```


### Figure 6C: Volcano plots showing genes Lungs over Primary Tumor in StealTHY KO
Colored points with FDR ≤ 0.10
```{r volcano-plot-stealthy-ko-pt-over-lung, fig.width=2, fig.asp=0.818}
lfc_colors <- c(none = "gray80", up = "#e41a1c", down = "#377eb8")
i <- 'stealthy_hunsg'

x <- gene_summ[[i]] %>% 
  ReadRRA(score = 'lfc') %>% 
  dplyr::rename(LFC = Score) %>% 
  mutate(
    direction = ifelse(FDR <= 0.10 & LFC > 0, 'up', 'none'),
    direction = ifelse(FDR <= 0.10 & LFC < 0, 'down', direction),
    is.significant = FDR <= 0.10,
    id = ifelse(id == 'TNFRSF10B', 'TRAIL-R2', id)
  )
use_xlim <- xlim(-max(abs(x$LFC), na.rm = TRUE), max(abs(x$LFC), na.rm = TRUE))

genes_to_highlight <- c('CCR10', 'TRAIL-R2', 'CCL27', 'IGFBP1', 'ANGPT2', 'AMH', 'IL34',
                   'IFNA8', 'TGFB2', 'IL13RA2', 'IFNAR1', 'IL24', 'IFNLR1', 'PTEN', 'GDF9')
data_text <- x %>% 
  filter(id %in% genes_to_highlight)

ggplot(x, aes(LFC, -log10(FDR), color = direction)) +
  geom_hline(yintercept = -log10(0.10), linetype = 3) +
  geom_vline(xintercept = 0, linetype = 3) +
  geom_point(shape = 16, alpha = 0.6, size = 0.8) +
  scale_color_manual(values = lfc_colors) +
  geom_text_repel(
    data = data_text,
    aes(label = id),
    color = 'black',
    size = 0.4*geom_text_size,
    min.segment.length = 0,
    segment.size  = 0.2,
    box.padding = 0.1,
    max.overlaps = Inf, 
    fontface="italic"
  ) +
  use_xlim +
  labs(
    x = expression(paste("lo", g[2],"(Fold change)")), 
    y = expression(paste("-lo", g[10],"(FDR)"))
  ) +
  guides(color = 'none', size = 'none') +
  theme(
      axis.text = element_text(size=6), 
      axis.title = element_text(size = 6)
  )
```

### Figure 6D: Distribution of the fold-change of Lungs over Primary Tumor in huNSG and NSG host
```{r split-violinplot-selected-genes-ic-vs-id-curated, fig.width=3.555, fig.asp=1.4}
i <- names(gene_summ)[1]

curate_gene_sets <- c(
  'Immune system cytokines and chemokines',
  'TGF-β family related',
  'Positive controls'
) 

levels_gene_sets <- c(
  'Immune system cytokines\nand chemokines',
  'TGF-β family related',
  'Positive controls'
) 

curated_gene_list <- c('IL24', 'IL13RA2', 'IL33', 'IL17C', 'IL17R', 'CCL27', 
                       'GDF9', 'TGFB2', 'AMH', 
                       'PTEN', 'LKB1', 'MET', 'EGFR')

curated_gene_info <- gene_info %>% 
  filter(gene_set %in% curate_gene_sets) %>% 
  filter(gene_name %in% curated_gene_list) 


plot_data <- foreach(i=names(gene_summ), .combine = rbind) %do% {
  xsg <- sgrna_summ[[i]] %>% 
     select(Gene, sgrna, LFC, FDR) %>% 
     dplyr::rename(sgLFC = LFC, sgFDR = FDR)

  gene_summ[[i]] %>% 
    ReadRRA(score = 'lfc') %>% 
    dplyr::rename(LFC = Score) %>% 
    mutate(
      direction = ifelse(FDR <= 0.10 & LFC > 0, 'up', 'none'),
      direction = ifelse(FDR <= 0.10 & LFC < 0, 'down', direction),
      is.significant = FDR <= 0.10,
      comparison = i
    ) %>% 
    left_join(curated_gene_info, by = c('id' = 'id')) %>% 
    filter(!is.na(gene_set)) %>% 
    mutate(
      gene_name = factor(gene_name, levels = curated_gene_list),
      gene_set = ifelse(
        gene_set == 'Immune system cytokines and chemokines',
        'Immune system cytokines\nand chemokines',
        gene_set
      ),
      gene_set = factor(gene_set, levels_gene_sets)
      ) %>% 
    left_join(xsg, by = c('id' = 'Gene'), relationship = "many-to-many")
  
}

plot_data %<>% 
  mutate(
    mouse_model_short = case_match(comparison, 
                            'stealthy_hunsg' ~ 'huNSG', 
                            'stealthy_nsg' ~ 'NSG'),
    mouse_model_short = factor(mouse_model_short,
                               levels = c('NSG', 'huNSG'))
  )

xlim <- c(-ceiling(max(abs(plot_data$LFC), na.rm = TRUE)), 
          ceiling(max(abs(plot_data$LFC), na.rm = TRUE)))

plot_data %>% 
  ggplot(aes(y = sgLFC, x = gene_name, fill = mouse_model_short)) +
  geom_split_violin(
    aes(y = sgLFC, x = gene_name, fill = mouse_model_short),
    alpha = 0.8,
    draw_quantiles = 0.5,
    linewidth = one_pt/2,
    scale = 'width'
    ) +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = 3,linewidth = one_pt/2) +
  scale_x_discrete(expand = expansion(add = c(1, 1))) +
  scale_fill_manual(values = palette_hu_stealthy_ko) +
  ylim(xlim) +
  facet_grid(rows = vars(gene_set), scales = 'free_y', space = 'free_y')+
  theme_facet +
  theme(
    # axis.line = element_line(linewidth = one_pt/2, color = 'black'),
      axis.ticks = element_line(linewidth = one_pt/2, color = 'black'),
      panel.border = element_rect(linewidth = one_pt/2, color = 'black'),
      axis.text = element_text(size=8), 
      axis.title = element_text(size = 8),
      legend.title = element_text(size=8),
      legend.text = element_text(size=8),
      strip.text = element_text(size = 4)
  ) +
  labs(
    y = expression(paste("lo", g[2],"(Fold change)")),
    x = NULL
  )+
  guides(color = 'none', fill = 'none', size = 'none')
  
```

### Supplementary Figure 6D: Fold-change of Lungs over Primary Tumor in huNSG and NSG host
```{r dotplot-selected-genes-allconditions, fig.width=2.97, fig.asp=2.8}
i <- names(gene_summ)[1]
plot_data <- foreach(i=names(gene_summ), .combine = rbind) %do% {
  gene_summ[[i]] %>% 
    ReadRRA(score = 'lfc') %>% 
    dplyr::rename(LFC = Score) %>% 
    mutate(
      direction = ifelse(FDR <= 0.10 & LFC > 0, 'up', 'none'),
      direction = ifelse(FDR <= 0.10 & LFC < 0, 'down', direction),
      is.significant = FDR <= 0.10,
      comparison = i
    ) %>% 
    left_join(gene_info, by = c('id' = 'id')) %>% 
    filter(!is.na(gene_set)) %>% 
    mutate(
      gene_name = factor(gene_name, levels = gene_levels),
      gene_set = ifelse(
        gene_set == 'Immune system cytokines and chemokines',
        'Immune system cytokines\nand chemokines',
        gene_set
      ),
      gene_set = ifelse(
        gene_set == 'Interferone, TNF and danger/death-inducing signals',
        'Interferon, TNF and\ndanger/death-inducing\nsignals',
        gene_set
      ),
      gene_set = factor(gene_set, names(use_colors))
      )
}

plot_data %<>% 
  mutate(
    mouse_model_short = case_match(comparison, 
                            'stealthy_hunsg' ~ 'huNSG', 
                            'stealthy_nsg' ~ 'NSG'),
    mouse_model_short = factor(mouse_model_short,
                               levels = c('huNSG', 'NSG'))
  )

xlim <- c(-ceiling(max(abs(plot_data$LFC), na.rm = TRUE)), 
            ceiling(max(abs(plot_data$LFC), na.rm = TRUE)))

plot_data %>% 
  ggplot(aes(LFC, gene_name, color = gene_set, size = -log10(FDR))) +
    geom_point(aes(shape = gene_set)) +
    geom_vline(xintercept = 0, linetype = 3) +
    scale_y_discrete(expand = expansion(add = c(1, 1))) + # add fixed units around each facet to avoid lines over points
    scale_x_continuous(expand = expansion(add = c(1, 1))) + # add fixed units around each facet to avoid lines over points
    scale_color_manual(values = use_colors) +
    scale_size(range = c(0.5, 2.5), limits = c(0, 4)) +
    scale_shape_manual(values = use_shape) +
    xlim(xlim) +
    facet_grid(cols = vars(mouse_model_short), rows = vars(gene_set), scales = 'free_y', space = 'free_y') +
    theme_facet +
    theme(
      # axis.line = element_line(linewidth = one_pt/2, color = 'black'),
      axis.ticks = element_line(linewidth = one_pt/2, color = 'black'),
      panel.border = element_rect(linewidth = one_pt/2, color = 'black'),
      axis.text = element_text(size=6), 
      axis.title = element_text(size = 6),
      legend.title = element_text(size=6),
      legend.text = element_text(size=6),
      strip.text = element_text(size = 4)
    ) +
    labs(
      x = expression(paste("lo", g[2],"(Fold change)")),
      y = NULL,
      size = expression(paste("-lo", g[10],"(FDR)"))
    ) +
    guides(color = 'none', shape = 'none')
```

